---
title: "Final Project Draft"
author: "Aidan Godfrey, Irene Zhang, Benjamin Lewis "
date: "4/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

### What is the purpose of the project?
 

* <font size="3">  Our analysis will be heavily based on grouping the data and seeing if the person uses their dominant hand or non-dominant hand more often <font/>
*  <font size="3">  We are studying two cases of wether over ten (10) seperate mealtimes with Irene Zhang's parent, how often do they touch there face with their dominant and non-dominant hand? <font/>


#### Goal: We want to see for each person, what hand are they more likely to use to touch their face.
  

<hr/>
  
## Project outlook

* <font size="3"> We want to see what hand a person uses to touch their face more often <font/>
* <font size="3"> This can be important in terms of a virus spreading, in that if someone uses their dominant hand to use things, then they touch there faces, viruses may spread at higher rates <font/>
*  <font size="3"> We will look at the total times two memberes of Irene's Family touches there face
* <font size="3"> Group them together to make a statment of wether a person use their dominant. <font>

> Null Hypothesis: H0 -- There is no difference in the difference between the proportion of what hand is used to touch someone's face (p dominant == p non-dominant)


> Alternative Hypothesis: HA -- There is a difference in the proportion of what hand is used to touch someone's face (p dominant != p non-dominant).

#### Note: For this test, we want to answer they use one hand more often for each person.

<hr/>

## Alternative outlook

We might be looking at a analyzitation of who touches there face more often per mealtime, difference of two proprotions. 

Who uses dominant hand more often, the mother or father?

> Null Hypothesis: H0 -- There is no difference between the proportion of hand touches with the dominant hand between Irene's Mother and Father. (p fatherdomi == p motherdomi)

> Alternative Hypothesis: HA -- There is a difference between the proportion of hand touches with the dominant hand between Irene's Mother and Father.(p fatherdomi != p motherdomi)


#### Note: For this test, we want to answer if a parent uses there dominant hand more often than the other. I THINK WE SHOULD USE THE FIRST ONE.

<hr/>

#### Data Collection

The overall data collection was done with these guidlines to ensure the integrity of the study

* Irene counted how often her parents touched their face in each of there mealtimes
* We also categorize the reasons for them to touch their face to be: 
  + Directly touching face
  + Scratch unintentionally or intentionally
  + Using of paper tissue
  + Touching their hair near their face
  + DISCLAIMER: using a utensil and touching the face was not included
* Everytime someone touched their face Irene would keep track on there phone without the parents noticing
* Irene has done the effort to keep the expereiment blind to both of the participants, because if they were aware of such a study, their behaviors would be abnormal.
* The experiment was done across 10 seperate consecutive days

<hr/>


 FINISH THIS ONCE IRENE PUBSLISHEs THE DATA
## Data Analysis

```{r data wrangle}
dad = c()
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
